---
title: 'Problem Set #4'
author: "Céline Vendler"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      error = FALSE, 
                      warning = FALSE)

# Load library "gt" in order to be able to create and format the chart in Question 2
library(gt)
library(tidyverse)

# Read in the file "ps_4_elections-poll-nc09-3.csv" and store as new dataset called "orig_data"
orig_data <- read_csv("ps_4_elections-poll-nc09-3.csv")
```

## Question 1

```{r mad_libs}
# To fill in Mad Lib 1, create a new dataset called "mad_lib_1" in order to replace the "X" with the content of "mad_lib_1" dynamically using inline R code (here, the content is a singular number); begin with the "orig_data" dataset
mad_lib_1 <- orig_data %>% 
  # Filter the data to keep only observations for respondents who responded "Dem"
  filter(response == "Dem") %>% 
  # Count the number of observations in order to determine the number of respondents who supported the Democratic candidate
  count()

# To fill in Mad Lib 2, create a new dataset called "mad_lib_2" in order to replace the "X" with the content of "mad_lib_2" dynamically using inline R code (here, the content is a singular number); begin with the "orig_data" dataset
mad_lib_2 <- orig_data %>% 
  # Filter the data to keep only observations for respondents who responded either "Rep" or "Und"
  filter(response %in% c("Rep", "Und")) %>% 
  # Count the observations by response (count the number of respondents who responded "Rep" and the number of respondents who responded "Und")
  count(response) %>% 
  # Spread the data using the variable "response" (the values of which are "Rep" and "Und") as the key and using the variable "n" (the values of which are the counts from above) as the value so that the dataset has two columns--one for "Rep" and one for "Und"--and one row, within which are the corresponding counts for "Rep" and "Und" responses
  spread(response, n) %>% 
  # Calculate the difference ("diff") between the count for Rep and the count for Und, reflecting how many more respondents there were who favored the Republican candidate than who were undecided
  summarize(diff = Rep - Und)
  
# To fill in Mad Lib 3, create a new dataset called "mad_lib_3" in order to replace the "X" with the content of "mad_lib_3" dynmically using inline R code (here, the content is a singular number); begin with the "orig_data" dataset
mad_lib_3 <- orig_data %>% 
  # Filter the data to keep only observations where the variables "gender" and "gender_combined" have different values
  filter(gender != gender_combined) %>% 
  # Count the number of observations in order to determine the number of individuals for whom the variables "gender" and "gender_combined" have different values
  count()

# To fill in Mad Lib 4, create a new dataset called "mad_lib_4" in order to replace the "X" with the content of "mad_lib_4" dynmically using inline R code (here, the content is a singular number); begin with the "orig_data" dataset
mad_lib_4 <- orig_data %>% 
  # Filter the data to keep only observations for respondents listed as "White" under "race_eth" but not listed as "White" under "file_race_black"
  filter(race_eth == "White", 
         file_race_black != "White") %>% 
  # Count the number of observations in order to determine the number of respondents listed as "White" under "race_eth" who are not listed as "White" under "file_race_black"
  count()

# To fill in Mad Lib %, create a new dataset called "mad_lib_5" in order to replace the "X" with the content of "mad_lib_5" dynmically using inline R code (here, the content is a singular number); begin with the "orig_data" dataset
mad_lib_5 <- orig_data %>% 
  # Filter the data to keep only observations for respondents who responded either "Dem" or "Rep"
  filter(response %in% c("Dem", "Rep")) %>% 
  # Group the data by the variable "response" in order to be able to compare first-response-timestamps across responses (here, "Dem" and "Rep")
  group_by(response) %>% 
  # Summarize the data in order to determine the timestamps of the first responses of "Dem" and "Rep" respectively ("first_response")
  summarize(first_response = min(timestamp)) %>% 
  # Spread the data using the varible "response" (the values of which are "Dem" and "Rep") as the key and using the variable "first_response" (the values of which are the timestamps of the first responses of "Dem" and "Rep" respectively) as the value so that the dataset has two columns--one for "Dem" and one for "Rep"--and one row, within which are the timestamp of the first "Dem" response and that of the first "Rep" response
  spread(response, first_response) %>% 
  # Calculate the time difference ("diff"), rounded to the nearest minute, between the timestamp of the first "Rep" response and that of the first "Dem" response in order to determine by how many minutes the first "Dem" response came before the first "Rep" response
  summarize(diff = round(Rep - Dem))
```

There were `r mad_lib_1` respondents who supported the Democratic candidate.

There were `r mad_lib_2` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (`gender` and `gender_combined`). There are `r mad_lib_3` individuals for whom these variables have different values.

There are `r mad_lib_4` respondents listed as “White” under `race_eth` who are not listed as “White” under  `file_race_black`.

The first `response` of Dem came `r mad_lib_5` minutes (rounded to the nearest minute) before the first `response` of Rep.

## Question 2

```{r create_chart}
# Create a new dataset called "data_to_chart" to chart using gt; begin with the "orig_data" dataset
data_to_chart <- orig_data %>% 
  # Select only the variables pertinent to the chart, which are "response", "race_eth", and "final_weight"
  select(response, race_eth, final_weight) %>% 
  # Filter the data so that answers to "race_eth" that read "[DO NOT READ] Don't know/Refused" are excluded, per the instruction within the brackets: "DO NOT READ"
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>% 
  # Reorder the levels of the factor "race_eth" so that "White" is first, "Black" is second, "Hispanic" is third, "Asian" is fourth, and "Other" is fifth
  mutate(race_eth = fct_relevel(race_eth, c("White", 
                                            "Black", 
                                            "Hispanic", 
                                            "Asian", 
                                            "Other"))) %>% 
  # Group the data by the variables "race_eth" and "response" in order to be able to calculate weighted totals (the sums of final weights) by combination of "race_eth" and "response"
  group_by(race_eth, response) %>% 
  # Calculate "race_eth" and "response" combinations' respective weighted totals ("weighted_total") by summing the final weights within those combinations
  summarize(weighted_total = sum(final_weight)) %>% 
  # Spread the data using the varible "response" (the values of which are "3", "Dem", "Rep", and "Und") as the key and using the variable "weighted_total" (the values of which are the weighted totals of the "race_eth" and "response" combinations, respectively) as the value so that the dataset has five columns--"race_eth", "3", "Dem", "Rep", and "Und"--and five rows, within which are the values of "race_eth"--"White", "Black", "Hispanic", "Asian", and "Other"--and the weighted totals by "race_eth" and "response"; replace values of "NA" with 0
  spread(response, weighted_total, 
         fill = 0) %>% 
  # Ungroup the data in order to be able to manipulate individual (ungrouped) observations
  ungroup() %>% 
  # Create a new variable called "total_by_race_eth", the values of which are the sums of the values of the columns "3", "Dem", "Rep", and "Und" by "race_eth" (the sums of the weighted totals by "race_eth" and "response" for each "race_eth")
  mutate(total_by_race_eth = `3` + Dem + Rep + Und) %>% 
  # Alter the variable "Dem" by making its values equal to the original values of "Dem" divided by the corresponding values of "total_by_race_eth" so that the values of "Dem" are now decimals, reflecting the proportions of White, Black, Hispanic, Asian, and Other respondents who responded "Dem", respectively
  mutate(Dem = Dem / total_by_race_eth) %>% 
  # Alter the variable "Rep" by making its values equal to the original values of "Rep" divided by the values of "total_by_race_eth" so that the values of "Rep" are now decimals, reflecting the proportions of White, Black, Hispanic, Asian, and Other respondents who responded "Rep", respectively
  mutate(Rep = Rep / total_by_race_eth) %>% 
  # Alter the variable "Und" by making its values equal to the original values of "Und" divided by the corresponding values of "total_by_race_eth" so that the values of "Und" are now decimals, reflecting the proportions of White, Black, Hispanic, Asian, and Other respondents who responded "Und", respectively
  mutate(Und = Und / total_by_race_eth) %>% 
  # Deselect the variables "total_by_race_eth" and "3", as they are not needed anymore
  select(-c(total_by_race_eth, `3`)) %>% 
  # Set all values of 0 to "NA"
  na_if(0)

# Chart "data_to_chart" using functions from the library "gt"
gt(data_to_chart) %>% 
  # Give the chart a title; I chose to make my title "Polling Results from 3rd Wave for North Carolina's 9th Congressional District," as the data used is polling data from 3rd wave for North Carolina’s 9th Congressional District
  tab_header(title = "Polling Results from 3rd Wave for North Carolina's 9th Congressional District") %>% 
  # Give the chart a source note, citing The New York Times Upshot and Siena College and their 2018 live polls; I chose the source provided in the violin plot from the problem set
  tab_source_note(source_note = "New York Times Upshot/Siena College 2018 live polls") %>% 
  # Relabel the chart's columns so that the label "race_eth" is removed, so that Dem reads "DEM.", so that Rep reads "REP.", and so that Und reads "UND."
  cols_label(race_eth = "", 
             Dem = "DEM.", 
             Rep = "REP.", 
             Und = "UND.") %>% 
  # Reformat the values in the rows corresponding to the columns "DEM.", "REP.", and "UND." so that the decimal values appear as whole-number percentages
  fmt_percent(columns = vars(Dem, Rep, Und), 
              decimals = 0) %>% 
  # Reformat the value housed in the fourth row of the "UND." column, which reads "NA", as a missing value, so that "NA" is replaced with an em-dash
  fmt_missing(columns = vars(Und), rows = 4)
```

## Question 3

```{r create_graphic, fig.width = 7.1}
# Set the width of the figure to 7.1 so that my subtitle does not get cut off when I knit this Rmd file (thank you to Jenny for this clever solution)

# Create a new dataset called "data_to_plot" to plot using ggplot; begin with the "orig_data" dataset
data_to_plot <- orig_data %>% 
  # Select only the variables pertinent to the plot, which are "educ" and "final_weight"
  select(educ, final_weight) %>% 
  # Filter the data so that answers to "educ" that read "[DO NOT READ] Refused" are excluded, per the instruction within the brackets: "DO NOT READ"
  filter(educ != "[DO NOT READ] Refused") %>% 
  # Reorder the levels of the factor "educ" so that "Grade school" is first, "High school" is second, "Some college or trade school" is third, "Bachelors' degree" is fourth, and "graduate or Professional Degree" is fifth
  mutate(educ = fct_relevel(educ, c("Grade school", 
                                    "High school", 
                                    "Some college or trade school", 
                                    "Bachelors' degree", 
                                    "Graduate or Professional Degree")))

# Plot "data_to_plot"; set "educ" as the x-axis variable and "final_weight" as the y-axis variable; these assignments will later be flipped using the "coord_flip" function, but the ultimate goal is to be able to visualize the density of weights given to respondents of various levels of education and to be able to make comparisons across levels of education
ggplot(data_to_plot, aes(x = educ, y = final_weight)) + 
  # Make the plot a violin plot so that the above goal can be accomplished
  geom_violin() + 
  # Flip the coordinates so that x becomes y and y becomes x; visually, "final_weight" appears as the x-axis variable, and "educ" appears as the y-axis variable
  coord_flip() + 
  # Add "jittered" points to the plot (add random variation to the locations of points); set height (which, because of the flipping of coordinates, really affects width) and width (which, for the same reason, really affects height) values for the random noise so that the points match those on the plot provided in the problem set as closely as possible; set alpha to 0.5 so that the points are semi-transparent (so that individual points appear grey and not black)
  geom_jitter(height = 0.1, width = 0.2, alpha = 0.5) + 
  # Remove the x-axis label (which, visually, appears as the y-axis label, so I am removing the label for all the educational levels on the plot); give the y-axis (which, visually, appears as the x-axis) the label "Weight Given to Respondent in Calculating Poll Results"; give the plot a title, subtitle, and caption to match the plot provided in the problem set
  labs(x = "", 
       y = "Weight Given to Respondent in Calculating Poll Results", 
       title = "More Educated Matter Less in North Carolina 9th", 
       subtitle = "Poll gives more weight to people who are less likely to participate in polls", 
       caption = "New York Times Upshot/Siena College 2018 live polls")
```

## Question 4

```{r create_my_plot, fig.width = 7.6}
# Set the width of the figure to 7.6 so that my subtitle does not get cut off when I knit this Rmd file

# Create a new dataset called "female_data" in order to manipulate only data for female respondents; begin with the "orig_data" dataset; I separate female and male data in this way because I encountered trouble when borrowing from the technique I used in Question 2 and, with help from my brief conversation with Albert after class, found that spreading the data, in this case, is not helpful
female_data <- orig_data %>% 
  # Filter the data to keep only observations for female respondents
  filter(gender == "Female") %>% 
  # Select only the variables pertinent to my plot, which I intend to be a stacked bar chart (from top to bottom, in decreasing order of percentages), with one bar reflecting the percentages of female respondents who responded "Dem", "Rep", "Und", and "3", respectively, and with the other bar reflecting the percentages of male respondents who responded "Dem", "Rep", "Und", and "3", respectively so that comparisons across the two genders can be made; the pertinent variables are thus "gender" and "response" (while it may seem redundant to select and group by "gender" in addition to "response", I need to so that female and male data can be differentiated from one another once I bind the rows of the two datasets, female and male)
  select(gender, response) %>% 
  # Group the data by the variables "gender" and "response" in order to be able to calculate totals (total number of respondents) by combination of "gender" and "response" ("total_by_combination")
  group_by(gender, response) %>% 
  # Calculate "gender" and "response" combinations' respective totals by counting the number of observations for those combinations; I do not, as I do in Question 2, use weighted values in this calculation, as the instructions of the problem set specify the variables we should use in this question, and "final_weight" is not one of them
  summarize(total_by_combination = n()) %>% 
  # Create a new variable called "total", the values of which are the sums of all the totals by combination, equal to the number of female respondents
  mutate(total = sum(total_by_combination)) %>% 
  # Create a new variable called "proportion", the values of which are the totals by combination divided by the sum of all the totals by combination, reflecting the proportions of female respondents who responded "Dem", "Rep", "Und", and "3", respectively
  mutate(proportion = total_by_combination / total)

# Create a new dataset called "male_data" in order to manipulate only data for male respondents; begin with the "orig_data" dataset
male_data <- orig_data %>% 
  # Filter the data to keep only observations for male respondents
  filter(gender == "Male") %>% 
  # Select only the variables pertinent to my plot, which are "gender" and "response"
  select(gender, response) %>% 
  # Group the data by the variables "gender" and "response" in order to be able to calculate totals (total number of respondents) by combination of "gender" and "response" ("total_by_combination")
  group_by(gender, response) %>% 
  # Calculate "gender" and "response" combinations' respective totals by counting the number of observations for those combinations
  summarize(total_by_combination = n()) %>% 
  # Create a new variable called "total", the values of which are the sums of all the totals by combination, equal to the number of male respondents
  mutate(total = sum(total_by_combination)) %>% 
  # Create a new variable called "proportion", the values of which are the totals by combination divided by the sum of all the totals by combination, reflecting the proportions of male respondents who responded "Dem", "Rep", "Und", and "3", respectively
  mutate(proportion = total_by_combination / total)

# Create a new dataset called "both" in order to combine the "female_data" and the "male_data" datasets into one dataset so as to be able to plot in one graphic both female and male data using ggplot; begin by binding together the rows of "female_data" and those of "male_data"
both <- bind_rows(female_data, male_data) %>% 
  # Deselect the variables "total_by_combination" and "total", as they are not needed anymore
  select(-c(total_by_combination, total)) %>% 
  # Reorder the levels of the factor "response" so that "Dem" is first, "Rep" is second, "Und" is third, and "3" is fourth
  mutate(response = fct_relevel(response, c("Dem", "Rep", "Und", "3")))

# Plot "both", setting "gender" as the x-axis variable, "proportion" as the y-axis variable (through the textual annotations below, I will convert the values of "proportion" from decimals to percentages below so that, visually, the unit of the y-axis is percentage, making the graphic easier to digest)
ggplot(both, aes(x = gender, y = proportion, fill = response)) + 
  # Make the plot a bar chart, specifically a "geom_col" chart
  geom_col() + 
  # Add textual annotations to the bars, labeling each stacked segment with its corresponding percentage, rounded to the nearest whole number (to reach these percentages, I multiply the values of "proportion" by 100 and round), pasting percent signs ("%") directly to the right of each percentage, separating them by no spaces; position these textual annotations so that they are stacked on top of each other as the individual segments of the bars are and so that each percentage appears in the middle of its corresponding segment; adjust the size of the textual annotations so that they both are visible and fit reasonably within their corresponding segments
  geom_text(aes(label = paste(round(100 * proportion), "%", sep = "")), 
            position = position_stack(vjust = 0.5), 
            size = 3) + 
  # Remove the x- and y-axis labels, as it is evident that the x-axis variable is "gender" from the labeling of "Female" and "Male" bars and as it is evident that the y-axis values correspond to percentages from the textual annotations; give the plot a title, subtitle, and caption indicating a source; in my subtitle, I chose to highlight the observations that the percentage of female respondents who responded "Dem" is equal to the percentage of male respondents who responded "Rep" and that, similarly, the percentage of female respondents who responded "Rep" is equal to the percentage of male respondents who responded "Dem"; I chose the source provided in the violin plot from the problem set
  labs(x = "", 
       y = "", 
       title = "Comparing Polling Results across Genders", 
       subtitle = "Female and male respondents support the Democratic and Republican candidates in opposite percentages", 
       caption = "New York Times Upshot/Siena College 2018 live polls") + 
  # Remove the y-axis tick labels; the y-axis ticks; and the title of the legend from the plot
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        legend.title = element_blank()) + 
  # Create my own discrete scale, relabeling levels "Dem" to "Democrat", "Rep" to "Republican", "Und" to "Undecided", and "3" to "Third Party"; assign aesthetic values so that "skyblue2" corresponds to "Democrat", "lightcoral" corresponds to "Republican", "thistle4" corresponds to "Undecided", and "plum3" corresponds to "Third Party"; I chose these colors intentionally, as blue is typically associated with the Democratic Party; as red is typically associated with the Republican Party; as grey indicates a "grey area," so to speak, and so corresponds with "Undecided"; and as purple is both off-blue and off-red and so corresponds with "Third Party"; I understand that the colors I chose are not perfectly blue, red, grey, and purple, but I chose colors I thought went nicely together guided by that color palette
  scale_fill_manual(labels = c("Democrat", "Republican", "Undecided", "Third Party"), 
                    values = c("skyblue2", "lightcoral", "thistle4", "plum3"))
```

## Colleagues

I attended Dillon's study hall, used Jenny's response to my Piazza question, and reached out to Albert for help with Question 4 and GitHub.
